Stochastic Approximation Monte Carlo with Dynamic Update Factor (SAD)

### 2018-08-08 papers/histogram/sad.tex
added some text on SAD

>>>>

{\color{red}
It is for this reason that we propose a way to determine the update factor
``dynamically'' while still satisfying the criteria for convergence.
For the update factor to scale well across both discrete and continuous systems,
it must include the number of states $N_S$. We define the average entropy as
$\bar S$ and the time since the simulation last found a new energy state $t_L$.
We then can write the following expression for the \emph{proposed} update factor.
\begin{align}
\gamma = N_S \frac{\bar S + \nicefrac{t}{t_L}}{N_{S}^2 + \nicefrac{t^2}{t_L}}
\end{align}
We now consider the following limiting cases:
\subsubsection{$t \propto N_S \approx t_L$ and $ t << N_S$}
In this limiting case, we find that the update factor reduces to
\begin{align}
\gamma \approx \frac{\bar S}{N_S} \approx \frac{\bar S}{t} 
\end{align}
Because $t \propto N_S$, the update factor scales as 1/t which is consistent
with the requirement for SAMC convergence.
\subsubsection{$t >> N_{S}^2$ and $ t\approx t_L$}
In this limiting case, we find that the update factor reduces to
\begin{align}
\gamma \approx \frac{N_S \bar S t_L}{t^2} \approx \frac{N_S \bar S}{t} 
\end{align}
\subsubsection{$t >> t_L \bar S$}
In this limiting case, we find that the update factor reduces to
\begin{align}
\gamma \approx \frac{N_S}{t}
\end{align}
We now integrate (sum) the update factor from the last time an energy state was found
to the final simulation time to ensure that convergence is guaranteed.
\begin{align}
\int_{t_L}^{t_f} \gamma dt \approx t_L N_S \left[ \int_{t_L}^{t_f} 
\frac{\bar S -1}{t^2} dt + \frac{1}{t_L}\int_{t_L}^{t_f} \frac{1}{t} dt \right]
\end{align}
From this we obtain
\begin{align}
\int_{t_L}^{t_f} \gamma dt \approx t_L N_S \left[\left(\bar S -1\right)\left(\frac{1}{t_L}-\frac{1}{t_f}\right) + \frac{1}{t_L} \ln\left(\frac{t_f}{t_L}\right) \right]
\end{align}
Simplifying, we see that as $t_f \rightarrow \infty$ then $\gamma \rightarrow \infty$ satisfying
the requirement for SAMC convergence.
\begin{align}
\gamma \approx \left(1-\frac{t_L}{t_f}\right)N_S\left(\bar S -1\right) + N_S \ln\left(\frac{t_f}{t_L}\right)
\end{align}
}
